\input{beamer_preamble.tex}
\begin{document}

\title[Virtual Large Cardinals]{Virtual Large Cardinals\\ {\small\textsc{British Logic Colloquium 2019, Oxford}}}
\author[Dan Saattrup Nielsen]{Dan Saattrup Nielsen\\ University of Bristol}
\date{September 5, 2019}

\begin{frame}
	\titlepage
\end{frame}

\begin{frame}{What is a large cardinal?}
  There is no rigorous definition of a large cardinal\pause, but many of them, especially in the higher reaches of the hierarchy, are characterised as the \textit{critical point} of a certain elementary embedding:
  \eq{
    \pi\colon(\M,\in)\to(\N,\in)
  }

  \pause
  \begin{block}{Examples}
    \begin{itemize}
      \item $\kappa$ is a \textbf{measurable cardinal} if $\M=H_{\kappa^+}$.
      \item $\kappa$ is a \textbf{$\theta$-strong cardinal} if $\M=H_\theta$, $H_\theta\subset\N$ and $\pi(\kappa)>\theta$.
    \end{itemize}
  \end{block}

  Recall that $x\in H_\theta$ iff $\abs{\tc(x)} < \theta$. This hierarchy is often more convenient than the $V_\alpha$'s since $H_\theta\models\zfc^-$ if $\theta$ is regular.
\end{frame}

{\setbeamercolor{background canvas}{bg = white}
\begin{frame}{The hierarchy of large cardinals}
  \pix[0.65]{gfx/hierarchy.pdf}
\end{frame}}

\begin{frame}{What is a \emph{virtual} large cardinal?}
  We basically just require that the embeddings exist in a \emph{generic extension} rather than in $V$:

  \pause

  \begin{block}{"Definition"}
    Let $\Phi$ be a large cardinal concept defined via elementary embeddings between \textit{sets}, like the definitions on the previous slide.

    \pause

    Then $\kappa$ is \textbf{virtually $\Phi$} if the same definition holds but where we only require the embeddings exist in a generic extension and that $\N\subset V$.
  \end{block}
\end{frame}

{\setbeamercolor{background canvas}{bg = white}
\begin{frame}{A virtual addition to the hierarchy}
  \pix[0.65]{gfx/hierarchy_w_virtuals.pdf}
\end{frame}}

\begin{frame}{Attaching an adjective}
  \only<1-3>{
    Let us attach a \textbf{pre-} to our large cardinals if we do not require anything about where the critical point is sent:

    \begin{block}{Example}
      $\kappa$ is \textbf{prestrong} if for every regular $\theta>\kappa$ there is an elementary embedding $\pi\colon (H_\theta,\in)\to(\N,\in)$ with $\crit\pi=\kappa$ and $H_\theta\subset\N$.
    \end{block}
  }

  \only<2>{  
    This is really not an interesting concept in the real world:

    \begin{block}{Proposition (folklore)}
      For regular cardinals $\kappa<\theta$:
      \begin{itemize}
        \item $\kappa$ is $\theta$-prestrong iff it is $\theta$-strong
        \item $\kappa$ is $\theta$-presupercompact iff it is $\theta$-supercompact
      \end{itemize}
    \end{block}
  }

  \only<3>{
    It \textit{is} interesting in the virtual world, however:

    \begin{block}{Theorem (N.)}
      For regular cardinals $\kappa<\theta$, $\kappa$ is virtually $\theta$-prestrong iff either
      \begin{itemize}
        \item $\kappa$ is virtually $\theta$-strong, or
        \item $\kappa$ is virtually $(\theta,\omega)$-superstrong
      \end{itemize}
    \end{block}
  }
\end{frame}

\begin{frame}{Characterising a phenomenon}
  \begin{block}{Corollary (N.)}
    Virtually $\theta$-prestrongs are equiconsistent with virtually $\theta$-strongs.
  \end{block}

  \pause

  \begin{block}{Corollary (N.)}
    The following are equivalent:
    \begin{itemize}
      \item Virtually prestrongs are equivalent to virtually strongs
      \item There are no virtually $\omega$-superstrongs.
    \end{itemize}
  \end{block}

  \pause

  Note that $\omega$-superstrong cardinals are inconsistent with $\zfc$!
\end{frame}

\begin{frame}{Adding parameters}
  \begin{block}{Definition}
    Let $\kappa<\theta$ be regular and let $A$ be a class. Then $\kappa$ is \textbf{virtually $(\theta,A)$-prestrong} if there exists a generic elementary embedding
    \eq{
      \pi\colon(H_\theta^V,\in,A\cap H_\theta^V)\to(\N,\in,B)
    }

    such that $\crit\pi=\kappa$, $H_\theta^V\subset\N$, $\N\subset V$ and $A\cap H_\theta^V=B\cap H_\theta^V$.

    \qquad Further, if $\pi(\kappa)>\theta$ then $\kappa$ is \textbf{virtually $(\theta,A)$-strong}.
  \end{block}

  \pause
  
  Can we find some virtual large cardinal characterising exactly when the virtually $A$-prestrongs are equivalent to virtually $A$-strongs?

  \pause

  Remember that we are looking for a large cardinal notion which is inconsistent in the real world.
\end{frame}

\begin{frame}{Berkeley cardinals}
  \begin{block}{Definition}
    $\delta$ is \textbf{virtually berkeley} if for every transitive set $\M$ there exists a generic elementary embedding $\pi\colon\M\to\M$ with $\crit\pi<\delta$.
  \end{block}

  \only<2>{The real world versions of these are of course inconsistent with $\zfc$, but are currently being investigated in a choiceless context.}

  \only<3->{
    \begin{block}{Definition}
      Say that $\on$ is \textbf{virtually (pre)woodin} if for every class $A$ there exists a virtually $A$-(pre)strong cardinal $\kappa$.
    \end{block}
  }

  \onslide<4->{
    \begin{block}{Theorem (N.)}
      The following are equivalent:
      \begin{itemize}
        \item $\on$ is virtually prewoodin iff it is virtually woodin
        \item There are no virtually berkeley cardinals
      \end{itemize}
    \end{block}
  }
\end{frame}

\begin{frame}{Open questions}
  \begin{block}{Question}
    Is the existence of a virtually berkeley cardinal equivalent to the statement that, for every class $A$, every virtually $A$-prestrong is virtually $A$-strong?
  \end{block}

  \begin{block}{Question}
    Are virtually $\omega$-superstrongs equivalent to virtually berkeleys?

    Wilson (2018) has shown this to be true in $L$.
  \end{block}
\end{frame}

{\setbeamercolor{background canvas}{bg = white}
\begin{frame}
  \phantom{hidden text}
  \begin{center}
    {\LARGE Thank you for your attention.}
  \end{center}

  \pix[0.65]{gfx/virtual_hierarchy.pdf}
\end{frame}}

\end{document}
